---
title: "For or Against? A Breakdown of Campaign Spending"
author: "Emma Livingston, Clara Rosenberg, and Kara VanAllen"
date: "October 31, 2017"
output: html_document
---

## Context and Background

We took information from the Federal Election Committee (FEC) to look at the ratios between contributions for and against candidates. The FEC provides information every election cycle on the candidates, the contributions of committees, and the committees making contributions. Looking at the 2011-2012  general election cycle for the House of Representatives, we looked at the contributions for and against the candidates. Since some transactions are for a certain candidate and other transactions are against a certain candidate, we wanted to know what the ratio was between money given for the election of the winning candidate, in contrast with the money given against the election of the losing candidate. Essentially, if any committee wants Joe Smith to win in a certain district, what is the winning combination of donations in support of Joe Smith and against Smith's opponent? 

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, include = FALSE, message = FALSE}
library(tidyverse)
```

```{r}
general_house <- house_elections %>%
  filter(general_votes != 0) #filter the house_elections data down to only the general elections
num_candidates <- general_house %>%
  arrange(state, district) %>%
  group_by(state, district) %>%
  summarize(N = n()) %>% #number of candidates in a district
  filter(N == 1) #1 candidate means there was only one candidate running in that district
#num_candidates looks at how many candidates ran unopposed in the house election. Since there are only eight who ran unoppoosed, and we are looking primarily at California, we decided that we did not need to make a change to the data based on that information. 
```

```{r}
house_candidates <- filter(candidates, cand_office == "H")
#filter so that the candidates table only includes candidates running for the house of representatives
```

```{r}
small_house <- general_house %>%
  full_join(house_candidates, by = c("fec_id" = "cand_id")) %>%
  select(fec_id, state, district, incumbent, candidate_name, party, general_votes, ge_winner)
#join general_house and house_candidates to connect information about election results
```

```{r}
small_house_contributions <- small_house %>%
  full_join(contributions, by = c("fec_id" = "cand_id")) %>%
  select(fec_id, state.x, district, incumbent, candidate_name, party, general_votes, ge_winner, transaction_type, transaction_amt)
#join small_house and contributions to connect election results, candidates, and contributions
```

```{r}
ca_house_contributions <- small_house_contributions %>%
  filter(state.x == "CA", transaction_type == "24A" | transaction_type == "24E")
#bring the house contributions table down to California, and only transactions of type 24A and 24E
```

```{r}
total_amount_tbl <- ca_house_contributions %>%
  group_by(candidate_name, transaction_type) %>%
  summarize(total_amount = sum(transaction_amt)) %>%
  filter(!is.na(transaction_type))
#condenses ca_house_contributions so that there is one row per candidate per contribution type
```

```{r}
big_ca_contributions <- ca_house_contributions %>%
  full_join(total_amount_tbl, by = c("candidate_name" = "candidate_name")) %>%
  select(fec_id, state.x, district, incumbent, candidate_name, party, general_votes, ge_winner, transaction_type.y, total_amount) %>%
  arrange(district)
#joins ca_house_contributions with total_amount_tbl to get the information of the former with the total from the latter
```

```{r}
small_ca_contributions <- unique(big_ca_contributions)
#gets rid of duplicates from big_ca_contributions
```

```{r}
contributions_eliminate <- small_ca_contributions %>%
  filter((ge_winner == "W" & transaction_type.y == "24E") | (ge_winner == "N" & transaction_type.y == "24A"))
#selects only 24E contributions for winners and 24A contributions for losers
```

```{r}
districts_eliminate <- contributions_eliminate %>%
  group_by(district) %>%
  summarize(N = n()) %>%
  filter(N == 2)
ca_districts <- districts_eliminate %>%
  inner_join(contributions_eliminate, by = c("district", "district"))
ca_districts$transaction_type.y <- factor(ca_districts$transaction_type.y)
  #includes only districts that have both 24E for winners and 24A for losers
```

```{r}
ggplot(data = ca_districts, aes(x = district, y = total_amount, fill = transaction_type.y)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(name = "Transaction Type", labels = c('money spent against the loser', 'money spent to support the winner'), values=c("#d01c8b", "#4dac26")) +
  ylab("Money Spent by Percent") +
  xlab("Districts in CA")
#plot the data so that each district's bar is the same height and the division between the two colors represents the proportion for each of the two categories
```

24A = % of money spent against the loser
24E = %  of money spent for the winner


```{r}
final_plot <- function(state_arg) {
  state_house_contributions <- small_house_contributions %>%
  filter(state.x == state_arg, transaction_type == "24A" | transaction_type == "24E")
  total_amount_any_state <- state_house_contributions %>%
      group_by(candidate_name, transaction_type) %>%
      summarize(total_amount = sum(transaction_amt))
  big_any_state_contributions <- state_house_contributions %>%
  full_join(total_amount_any_state, by = c("candidate_name" = "candidate_name")) %>%
  select(fec_id, state.x, district, incumbent, candidate_name, party, general_votes, ge_winner, transaction_type.y, total_amount) %>%
  arrange(district)
  small_any_state_contributions <- unique(big_any_state_contributions)
  contributions_eliminate_function <- small_any_state_contributions %>%
  filter((ge_winner == "W" & transaction_type.y == "24E") | (ge_winner == "N" & transaction_type.y == "24A"))
  districts_eliminate_function <- contributions_eliminate_function %>%
  group_by(district) %>%
  summarize(N = n()) %>%
  filter(N == 2)
any_state_districts <- districts_eliminate_function %>%
  inner_join(contributions_eliminate_function, by = c("district", "district"))
any_state_districts$transaction_type.y <- factor(any_state_districts$transaction_type.y)
ggplot(data = any_state_districts, aes(x = district, y = total_amount, fill = transaction_type.y)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(name = "Transaction Type", labels = c('money spent against the loser', 'money spent to support the winner'), values=c("#d01c8b", "#4dac26")) +
  ylab("Money Spent by Percent") +
  xlab(paste("Districts in", state_arg))
}
#this function takes the process outlined above, starting when ca_house_contributions (the first state-specific table) was made, and applies it to any state. names of intermediate tables are changed so as to apply to all states and to avoid confusion.
```

```{r}
states <- c("AZ", "CA",  "CO", "CT", "FL", "GA", "HI", "IL", "IN", "IA", "KY", "LA", "MD", "MA", "MI", "MN", "NV", "NH", "NJ", "NY", "NC", "ND", "OH", "PA", "RI", "TX", "UT", "WA", "WV", "WI")
lapply(states, FUN = final_plot)
#this selects only the states that had at least one district with both 24E contributions for the winner and 24A contributions for the loser and applies the above function to them.
```


## Summary

For any state, this graphic takes information about the 2012 house elections and maps, per district, the percent of money, spent by any committee, in support of the candidate who won against the percent of money, spent by any committee, in opposition of the candidate who lost. Not all states and districts are included because not all districts had both 24E contributions for the winner and 24A contributions for the loser, and not every state has at least one such district. This limits the ability to apply these trends to all districts, but is a good graph for only considering these kinds of districts.
By using past trends to predict the future, we can learn where to spend money to make it most effective in getting the desired outcome.

Looking at these graphs, the takeaway is that almost all of the districts that meet the criteria spend most of their money in these two categories either in favor of the candidate who won, or in opposition to the candidate who lost. This means that the divide between money in support of the winner and money against the loser is not typically an even split. If we were to perform a statistical analysis of these data, we would likely see that it is more effective from a campaign finance standpoint to focus funds one way or the other to elect the candidate of choice.

For example, consider New York. Between the two categories of contribution in consideration, over 75% of money spent in each of those individual district campaigns were spend 'against the loser.' Clearly, the New York voters in those districts are more receptive to negative propaganda than they are to positive.