---
title: "Mini-Project 2"
author: "Kara VanAllen"
date: "October 31, 2017"
output: html_document
---

## Loading the data

This project is so interesting and relevant to my life!

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, include = FALSE, message = FALSE}
library(tidyverse)
```

```{r}
general_house <- house_elections %>%
  filter(general_votes != 0) #filter the house_elections data down to only the general elections
num_candidates <- general_house %>%
  arrange(state, district) %>%
  group_by(state, district) %>%
  summarize(N = n()) %>% #number of candidates in a district
  filter(N == 1) #1 candidate means there was only one candidate running in that district
# num_candidates looks at how many candidates ran unopposed in the house election. Since there are only eight who ran unoppoosed, and we are looking primarily at California, we decided that we did not need to make a change to the data based on that information. 
```

```{r}
house_candidates <- filter(candidates, cand_office == "H") #filter so that the candidates table only includes candidates running for the house of representatives
```

```{r}
small_house <- general_house %>%
  full_join(house_candidates, by = c("fec_id" = "cand_id")) %>%
  select(fec_id, state, district, incumbent, candidate_name, party, general_votes, ge_winner) #join general_house and house_candidates to connect information about election results
```

```{r}
small_house_contributions <- small_house %>%
  full_join(contributions, by = c("fec_id" = "cand_id")) %>%
  select(fec_id, state.x, district, incumbent, candidate_name, party, general_votes, ge_winner, transaction_type, transaction_amt) #join small_house and contributions to connect election results, candidates, and contributions
```

```{r}
ca_house_contributions <- small_house_contributions %>%
  filter(state.x == "CA", transaction_type == "24A" | transaction_type == "24E")
#bring the house contributions table down to California, and only transactions of type 24A and 24E
```

```{r}
total_amount_tbl <- ca_house_contributions %>%
  group_by(candidate_name, transaction_type) %>%
  summarize(total_amount = sum(transaction_amt)) %>%
  filter(!is.na(transaction_type)) #condenses ca_house_contributions so that there is one row per candidate per contribution type
```

```{r}
big_ca_contributions <- ca_house_contributions %>%
  full_join(total_amount_tbl, by = c("candidate_name" = "candidate_name")) %>%
  select(fec_id, state.x, district, incumbent, candidate_name, party, general_votes, ge_winner, transaction_type.y, total_amount) %>%
  arrange(district) #joins ca_house_contributions with total_amount_tbl to get the information of the former with the total from the latter
```

```{r}
small_ca_contributions <- unique(big_ca_contributions)
#gets rid of duplicates from big_ca_contributions
```

```{r}
contributions_eliminate <- small_ca_contributions %>%
  filter((ge_winner == "W" & transaction_type.y == "24E") | (ge_winner == "N" & transaction_type.y == "24A"))
#selects only 24E contributions for winners and 24A contributions for losers
```

```{r}
districts_eliminate <- contributions_eliminate %>%
  group_by(district) %>%
  summarize(N = n()) %>%
  filter(N == 2)
ca_districts <- districts_eliminate %>%
  inner_join(contributions_eliminate, by = c("district", "district"))
ca_districts$transaction_type.y <- factor(ca_districts$transaction_type.y)
  #includes only districts that have both 24E for winners and 24A for losers
```

```{r}
ggplot(data = ca_districts, aes(x = district, y = total_amount, fill = transaction_type.y)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_discrete(name = "Transaction Type", labels = c('money spent against the loser', 'money spent for the winner'))
```

<<<<<<< HEAD
24A = % of money spent against the loser
24E = %  of money spent for the winner

=======
```{r}
final_plot <- function(state_arg) {
  state_house_contributions <- small_house_contributions %>%
  filter(state.x == state_arg, transaction_type == "24A" | transaction_type == "24E")
  total_amount_any_state <- state_house_contributions %>%
      group_by(candidate_name, transaction_type) %>%
      summarize(total_amount = sum(transaction_amt))
  big_any_state_contributions <- state_house_contributions %>%
  full_join(total_amount_any_state, by = c("candidate_name" = "candidate_name")) %>%
  select(fec_id, state.x, district, incumbent, candidate_name, party, general_votes, ge_winner, transaction_type.y, total_amount) %>%
  arrange(district)
  small_any_state_contributions <- unique(big_any_state_contributions)
  contributions_eliminate_function <- small_any_state_contributions %>%
  filter((ge_winner == "W" & transaction_type.y == "24E") | (ge_winner == "N" & transaction_type.y == "24A"))
  districts_eliminate_function <- contributions_eliminate_function %>%
  group_by(district) %>%
  summarize(N = n()) %>%
  filter(N == 2)
any_state_districts <- districts_eliminate_function %>%
  inner_join(contributions_eliminate_function, by = c("district", "district"))
any_state_districts$transaction_type.y <- factor(any_state_districts$transaction_type.y)
ggplot(data = any_state_districts, aes(x = district, y = total_amount, fill = transaction_type.y)) +
  geom_bar(stat = "identity", position = "fill")
}
```

```{r}
final_plot("IN")
```
>>>>>>> 80402178463229fda08bd9ff364228033c8f167c

